---
title: "Assignment 2"
author: Group 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
embed-resources: true
editor: visual
execute:
  warning: false
  message: false
  error: false
  tidy: true
format: pdf
theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(writexl)
library(text)
library(readxl)
library(tidyverse)

```

We will explore sentiment analysis to give us an advantage in coding comments.

Read in data from github, and tidy it to continue processing.

```{r}

reddit_posts <- read_xlsx(
  "~/repos/SURV622_Assignment/data/posts_data_clean.xlsx") 

# comments data
reddit_comment <- read_xlsx(
  "~/repos/SURV622_Assignment/data/comments_data_clean.xlsx") 

```

We can use a model from huggingface to code each comment as positive, negative, or neutral.

We build a function that we will use to loop through comments and get scores.

```{r}
sentiment_score_fun <- function(comment_text){
  
  description = str_trunc(comment_text, 1600) 
  
  # shorten descriptions to confirm to the model limitation
  
  # use huggingface model to detect emotional tone of summary
  mod = 
    textClassify(description,
                 model_info = "ProsusAI/finbert", 
                 return_all_scores = FALSE, 
                 return_incorrect_results = FALSE,
                 function_to_apply = "softmax",
                 tokenizer_parallelism = TRUE,
                 device = "gpu") |>   # change to "cpu" if a gpu is not configured
    pull(1)
  
  return(mod)
  
}
```

Loop the comments through our function.

```{r}
reddit_comment_new <- reddit_comment |> 
  mutate(label = ifelse(is.na(comment), NA, sentiment_score_fun(comment)))

```

```{r}
 reddit_posts_new <- reddit_posts |> mutate(label = ifelse(is.na(text), NA, sentiment_score_fun(text)))
```

Add new column for hand coding (dis)approval

```{r}
# write posts data
write_xlsx(reddit_posts_new, "~/repos/SURV622_Assignment/data/posts_data_clean.xlsx")

# write comments data
write_xlsx(reddit_comment_new, "~/repos/SURV622_Assignment/data/comments_data_clean.xlsx")

```

Separate out data for hand coding based on positive and negative scores for keyword searches rif, reduction in force, and fork in the road to concentrate comments on the topic of federal layoffs. Sample 400 comments without replacement from the negative and positive coded comments.

```{r}
sample_comments_dat <- reddit_comment |> 
  filter(search_term %in% c("rif", "reduction in force", "fork in the road"),
         label %in% c("positive", "negative")) |> 
  #sample_n(size=400) |> 
  mutate(outcome = NA)

# group members
group_members <-c("Kevin", 'Felix', "Aria", "Gloria")

sample_comments_list <- infer::rep_slice_sample(sample_comments_dat, n=100, replace = FALSE, reps=4) |> 
  group_by(replicate) |> 
  group_split() 

walk(seq(1,4), function(x){
  
  write_xlsx(# pass data from list
    sample_comments_list[[x]] |> 
      select(-replicate), 
    # where to save, & how to name it
    str_c(
    "~/repos/SURV622_Assignment/data/comments_to_code/",
    group_members[x], ".xlsx")
  )
})

# write files to github by name

```
